---
output:
  pdf_document: default
  html_document: default
---

---
title: Bail Analysis 
author: |
  | Jamie Zeng
  | Strategic Planning & Policy
date: "Last Run: `r format(Sys.time(), '%d %B, %Y')`"
header-includes:
- \usepackage{geometry}
- \usepackage{caption}
- \usepackage{lscape}
- \usepackage{longtable}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
fontsize: 11pt
geometry: margin=1cm
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,warning=F,message=F,cache=F,fig.width=15,fig.height=7)
knitr::knit_hooks$set(inline = function(x) { prettyNum(x, big.mark=",") })

setwd("P:/Bureau Data/PLANNING/Management Reports/R Reports/Bail Analysis")

rm(list=ls())

library(RODBC);library(tidyverse);library(tidyr);library(dplyr);library(lubridate);library(janitor);library(forcats); library(kableExtra); library(scales);
library(stringr);  library(ggplot2);library(gt); library(stringr); library(gridExtra)

 dms <- odbcConnect('dms')

 data1 <- sqlQuery(dms, "SELECT * FROM ##viol_arr_citywide") %>%
   mutate(NYSID = str_squish(NYSID))


 data2 <- sqlQuery(dms, "SELECT * FROM ##dany_atliberty")  %>%
   mutate(NYSID = str_squish(NYSID))


 data3 <- sqlQuery(dms,"SELECT * FROM ##dany_vio_detail")  %>%
   mutate(NYSID = str_squish(NYSID))

 
# data1 <- read_rds('data1.RDS')
# data2 <- read_rds('data2.RDS')
# data3 <- read_rds('data3.RDS')  
   
## Number of arrests on office violent or violent felony offenses were made in Manhattan, involving how many persons. 
citywide_arr <- 
merge(
  data1 %>% 
  filter(ChargeOrder==1) %>% 
  mutate(offense_year = ifelse(offense_year == "2021", "2021 to date", offense_year)) %>% 
  mutate(`Year` = factor(offense_year, levels = c('2018', '2019', '2020', '2021 to date'))) %>% 
  group_by(`Year`) %>% 
  summarise(`VFOs` = n_distinct(ArrestID),
            `Persons Arrested on VFOs` = n_distinct(coalesce(NYSID, ArrestID))
 ),
 merge(data1 %>% 
        filter(ChargeOrder==1, OutInNY==1) %>% 
         select(NYSID, ArrestID, offense_date, offense_year),
       data2 %>% 
         filter(VioArrests > 0) %>% 
         select(NYSID, DefendantId, ArcDate, DispoDate),
       by = 'NYSID'
 ) %>% 
filter(offense_date >= ArcDate, offense_date <= coalesce(DispoDate, today())) %>% 
mutate(offense_year = ifelse(offense_year == "2021", "2021 to date", offense_year)) %>% 
group_by(`Year` = factor(offense_year, levels = c('2018', '2019', '2020', '2021 to date'))) %>% 
summarise(`VFOs by Persons at Liberty` = n_distinct(ArrestID),
          `Persons at Liberty Arrested on VFOs` = n_distinct(NYSID),
          `DANY Cases Associated with Persons at Liberty` = n_distinct(DefendantId))
, by = 'Year'
) %>% 
adorn_totals(where = 'row') %>% 
mutate(`% VFOs by Person at Liberty` = `VFOs by Persons at Liberty`/`VFOs`,
       `% Persons At Liberty of Persons Arrested on VFOs ` = `Persons at Liberty Arrested on VFOs`/`Persons Arrested on VFOs`) %>% 
mutate_at(vars(7:8), list(~percent(., accuracy = .01))) %>% 
select(1,2,4,7, 3,5,8, 6) 



man_arr <- 
merge(
  data1 %>% 
  filter(ArrestBorough == 'M', ChargeOrder==1) %>% 
  mutate(offense_year = ifelse(offense_year == "2021", "2021 to date", offense_year)) %>% 
  mutate(`Year` = factor(offense_year, levels = c('2018', '2019', '2020', '2021 to date'))) %>% 
  group_by(`Year`) %>% 
  summarise(`VFOs` = n_distinct(ArrestID),
            `Persons Arrested on VFOs` = n_distinct(coalesce(NYSID, ArrestID))
 ),
 merge(data1 %>% 
        filter(ArrestBorough == 'M', ChargeOrder==1, OutInNY==1) %>% 
         select(NYSID, ArrestID, offense_date, offense_year),
       data2 %>% 
         filter(VioArrests > 0) %>% 
         select(NYSID, DefendantId, ArcDate, DispoDate),
       by = 'NYSID'
 ) %>% 
filter(offense_date >= ArcDate, offense_date <= coalesce(DispoDate, today())) %>% 
mutate(offense_year = ifelse(offense_year == "2021", "2021 to date", offense_year)) %>% 
group_by(`Year` = factor(offense_year, levels = c('2018', '2019', '2020', '2021 to date'))) %>% 
summarise(`VFOs by Persons at Liberty` = n_distinct(ArrestID),
          `Persons at Liberty Arrested on VFOs` = n_distinct(NYSID),
          `DANY Cases Associated with Persons at Liberty` = n_distinct(DefendantId))
, by = 'Year'
) %>% 
adorn_totals(where = 'row') %>% 
mutate(`% VFOs by Person at Liberty` = `VFOs by Persons at Liberty`/`VFOs`,
       `% Persons At Liberty of Persons Arrested on VFOs ` = `Persons at Liberty Arrested on VFOs`/`Persons Arrested on VFOs`) %>% 
mutate_at(vars(7:8), list(~percent(., accuracy = .01))) %>% 
select(1,2,4,7, 3,5,8, 6) 

  
## X Number and Y percent of arrests during Z period occurred while the alleged person was at liberty on a pending DANY case, where the 
## person was ROR, Released under supervision or paid bail. 


# arc <- 
#   data2 %>% 
#   mutate(`Year` = factor(ArcYear, levels = c('2018', '2019', '2020', '2021'))) %>% 
#   group_by(`Year`) %>% 
#   summarise(`Cases Arraigned` = n_distinct(DefendantId),
#             `Violent Arrests` = sum(VioArrests), 
#             `persons Involved with Violent Arrests` = n_distinct(Nysid[VioArrests != 0]),
#             `Percentage of persons` = percent(n_distinct(Nysid[VioArrests != 0]) / n_distinct(Nysid), accuracy=0.01)) %>% 
#   adorn_totals("row") 


rs <-
  data2 %>% 
  mutate(`Release Status` = factor(ArcReleaseStatus, levels = c('Bail', 'Bail With Curfew', 'ROR', 'ROR With Curfew', 'Supervised Release', 'Intensive Community Monitoring','Parole'))) %>% 
  mutate(ArcYear = ifelse(ArcYear == "2021", "2021 to date", ArcYear)) %>% 
  mutate(`Year` = factor(`ArcYear`, levels = c('2018', '2019', '2020', '2021 to date'))) %>%
  group_by(`Year`,`Release Status`) %>% 
  summarise(`Cases Arraigned` = n_distinct(DefendantId)) %>% 
  mutate(`Percentage of Cases Arraigned` = percent(`Cases Arraigned`/ sum(`Cases Arraigned`), accuracy = 0.01))
# %>% 
#   adorn_totals("row")
  
summary(data1 %>% 
          filter(OutInNY==1, ChargeOrder==1))
## X Number of these arrests occurred while the alleged person was out on a misdemeanor, and Y Number while out on a violent felony. 
## Z of these arrests occurred while the alleged person was out on bail 



cat2 <- 
  data1 %>% 
  filter(OutInNY==1, ChargeOrder == 1) %>% 
  mutate(`Category` = factor(TopCatPendingDany, levels = c('Violent Felony',  'Non-Violent Felony', 'Misdemeanor', 'Violation/Infraction', 'Unknown'))) %>%
  mutate(offense_year = ifelse(offense_year == "2021", "2021 to date", offense_year)) %>% 
  mutate(`Year` = factor(`offense_year`, levels = c('2018', '2019', '2020', '2021 to date'))) %>% 
  group_by(`Year`,`Category`) %>% 
  summarise(`Number of Violent Arrests` = n_distinct(ArrestID)) %>% 
  mutate(`Percentage of Violent Arrests` = 
          percent(`Number of Violent Arrests` / sum(`Number of Violent Arrests`), accuracy=0.01)) %>% 
  adorn_totals("row") 



top2 <- 
     data1 %>% 
      filter(OutInNY==1, ChargeOrder == 1) %>% 
      mutate(`Top Charge` = paste0(TopChgPendingDany, ' - ', TopTxtPendingDany, ' (', TopCatPendingDany, ')')) %>% 
      group_by(offense_year = as.character(offense_year),`Top Charge`) %>% 
      summarise(`Number of Violent Arrests` = n_distinct(ArrestID)) %>%
      mutate(`Percentage of Violent Arrests` = 
          percent(`Number of Violent Arrests` / sum(`Number of Violent Arrests`), accuracy=0.01)) %>% 
      arrange(desc(`Number of Violent Arrests`)) %>% 
      top_n(5) %>% 
      group_by(offense_year) %>% 
      mutate(offense_year = ifelse(offense_year == "2021", "2021 to date", offense_year)) %>% 
      mutate( Rank = as.integer(rank(-`Number of Violent Arrests`))) %>% 
      select('Year' = offense_year, Rank, `Top Charge`, `Number of Violent Arrests`, `Percentage of Violent Arrests`) %>% 
      arrange(Year, Rank)



#####added section below

#persons who had 4 or more arrests while at liberty on a DANY case. 
arrPer <-
  data2 %>% 
  mutate(`Number of Violent Arrests Per Person` = VioArrests_Man) %>% 
  group_by(`Number of Violent Arrests Per Person`) %>% 
  summarise(`Number of Persons` = n_distinct(NYSID)) %>% 
  mutate(Percent = percent(`Number of Persons`/sum(`Number of Persons`), accuracy=0.01)) 


#plot not working?
ggplot(arrPer, aes(x = `Number of Violent Arrests Per Person`, y = `Number of Violent Arrests Per Person`)) +
  geom_line() +
  theme_bw() 
 

#top arrest charges for persons who had 4 or more arrests while at liberty on a DANY case.
topVio <- 
      data2 %>% 
      filter(VioArrests_Man >= 4)  
  

VioPer <- 
      data1 %>% 
      filter(NYSID %in% topVio$NYSID & ChargeOrder==1 & OutInNY==1) %>% 
      mutate(`Top Arrest Charge` = paste0(Charge, ' - ', ChargeDescription, ' (', ChargeCategory, ')')) %>% 
      group_by(offense_year = as.character(offense_year),`Top Arrest Charge`) %>% 
      summarise(`Number of Violent Arrests` = n_distinct(ArrestID)) %>%
      mutate(`Percentage of Violent Arrests` = 
          percent(`Number of Violent Arrests` / sum(`Number of Violent Arrests`), accuracy=0.01)) %>%
      arrange(desc(`Number of Violent Arrests`)) %>% 
      top_n(5) %>% 
      group_by(offense_year) %>% 
      mutate(offense_year = ifelse(offense_year == "2021", "2021 to date", offense_year)) %>% 
      mutate( Rank = as.integer(rank(-`Number of Violent Arrests`))) %>% 
      select('Year' = offense_year, Rank, `Top Arrest Charge`, `Number of Violent Arrests`, `Percentage of Violent Arrests`) %>% 
      arrange(Year, Rank)




 


#arrest alert for persons who had 4 or more arrests while at liberty on a DANY case.
arrestAlert <- 
      data3 %>% 
      mutate(`Arrest Alert` = ArrestAlert) %>% 
      group_by(`Arrest Alert`) %>% 
      summarise(`Number of Persons` = n_distinct(NYSID)) %>% 
      arrange(desc(`Number of Persons`))


#average conviction for persons who had 4 or more arrests while at liberty on a DANY case.     
avgCvt <-
      data3 %>% 
      distinct_at(vars(1,3,4,5,6)) %>% 
      mutate(`Average Conviction` = round(mean(`crimHis(Conviction)`)),
             `Average Felony Conviction` = round(mean(`crimHis(FelConv)`)),
             `Average Misdemeanor Conviction` = round(mean(`crimHis(MisdConv)`)),
             `Average Violation Conviction`= round(mean(`crimHis(ViolationConv)`))
               ) %>%    
      select(6,7,8,9) %>% 
      head(1)
      

#number of gun cases since 2018 for persons who had 4 or more arrests while at liberty on a DANY case.
gunCases <-
      data3 %>% 
      group_by(`Number of Gun Cases Since 2018` = GunCasesSince2018) %>% 
      summarise(`Number of Persons` = n_distinct(NYSID)) 
```



\blandscape
This analysis provides statistics on violent arrests since 2018, and statistics of those violent arrests made by persons who were at liberty following criminal court arraignment on a DANY case.




### Violent Arrests 

Below charts show numbers of arrests on office violent or violent felony offenses (VFOs) city wide vs. in Manhattan. Following fields are included:

- VFOs by Persons at Liberty: Number of VFOs committed by persons who were at liberty at the time of the offense.

- % VFOs by Persons at Liberty: Percentage of VFOs committed by persons who were at liberty at the time of the offense.

- Persons Arrested on VFOs: Number of persons who were arrested on VFOs.

- Persons at Liberty Arrested on VFOs: Number of persons who were arrested on VFOs while they were at liberty.

- % Persons at Liberty Arrested on VFOs: Percentage of persons who were arrested on VFOs while they were at liberty.

- DANY Cases Associated with Persons at Liberty: Number of DANY cases associated with persons who were arrested on VFOs while they were at liberty.

### City wide

```{r, echo = F}
citywide_arr %>% 
  kable(format = 'latex', digits = 0, longtable = T,
      booktabs = T, align = c(rep('l', 2), rep('c', 2)),
      format.args = list(big.mark = ","), linesep = "") %>% 
  column_spec(column = 2:8, width = "0.9in") %>% 
    kable_styling(font_size = 9, latex_options = "scale_down")
```
\elandscape

\blandscape
### Manhattan

```{r, echo = F}
man_arr %>% 
  kable(format = 'latex', digits = 0, longtable = T,
      booktabs = T, align = c(rep('l', 2), rep('c', 2)),
      format.args = list(big.mark = ","), linesep = "") %>% 
    column_spec(column = 2:8, width = "0.9in") %>% 
    kable_styling(font_size = 9, latex_options = "scale_down")
```
\elandscape


### Top Charge Category  

Below table shows the top charge category the alleged person was out on and the number of violent offenses committed by these persons. (If the person had multiple pending DANY cases while at liberty, they will be only counted once based on the most serious charge category across all cases the person had pending at the time of the offense.)


```{r, echo = F}
cat2 %>% 
kable(format = "latex",booktabs = T,
      align = c("l",rep("c",ncol(.))),digits = 0,
      format.args = list(big.mark = ","),linesep = "") %>%
  kable_styling(latex_options = c("HOLD_position")) %>% 
 column_spec(c(1:4), width = '4cm') %>% 
 collapse_rows(columns = 1, valign = 'top')
```


\blandscape

### Top Charge

Below table shows the top 5 charges the alleged person was out on and the number of violent offenses committed by these persons. please note that it only counts the top charge of the case. (If the person had multiple pending DANY cases while at liberty, they will be only counted once based on the most serious charge category across all cases the person had pending at the time of the offense.)

```{r, echo = F}
top2 %>% 
  kable(format = 'latex', digits = 0, longtable = T,
      booktabs = T, align = c(rep('l', 2), rep('c', 4)),
      format.args = list(big.mark = ","), linesep = "") %>% 
    kable_styling(font_size = 8, latex_options = c("HOLD_position")) %>% 
 collapse_rows(columns = 1, valign = 'top')

```
\elandscape

\blandscape

### Release Status at Arraignment  


Of those persons who committed violent offenses while at liberty, below chart provides the number and percentage of the release status that they were out on following arraignment. 


```{r, echo = F}

rs %>% 
kable(format = "latex",booktabs = T,
      align = c("l",rep("c",ncol(.))),digits = 0,
      format.args = list(big.mark = ","),linesep = "") %>%
 column_spec(c(1:4), width = '4cm') %>% 
 collapse_rows(columns = 1, valign = 'top') %>% 
    kable_styling(font_size = 8, latex_options = c("HOLD_position")) 

```
\elandscape


\blandscape

### Persons with Multiple Arrests While at Liberty on a DANY Case 

Of those who were at liberty on a DANY case, below chart and gragh provide distribution of number of violent arrests in Manhattan per person. 

```{r, echo = F}
arrPer %>% 
  kable(format = 'latex', digits = 0, longtable = T,
      booktabs = T, align = c(rep('l', 2), rep('c', 2)),
      format.args = list(big.mark = ","), linesep = "") %>% 
  column_spec(column = 2:8, width = "0.9in") %>% 
    kable_styling(font_size = 9, latex_options = "scale_down")
```
\elandscape


\blandscape

### Top Arrest Charge

Below table shows the top 5 arrest charges of persons with more than 4 violent arrests in Manhattan while at liberty on a DANY case. Please note that it only counts the top charge of the arrest


```{r, echo = F}
VioPer %>% 
  kable(format = 'latex', digits = 0, longtable = T,
      booktabs = T, align = c(rep('l', 2), rep('c', 4)),
      format.args = list(big.mark = ","), linesep = "") %>% 
     column_spec(3, width = '9.5cm') %>% 
    kable_styling(font_size = 8, latex_options = c("HOLD_position")) %>% 
 collapse_rows(columns = 1, valign = 'top')
```
\elandscape


\blandscape
### Arrest Alert 

Below table shows the arrest alerts of persons with more than 4 violent arrests in Manhattan while at liberty on a DANY case.

```{r, echo = F}
arrestAlert %>% 
  kable(format = 'latex', digits = 0, longtable = T,
      booktabs = T, align = c(rep('l', 2), rep('c', 2)),
      format.args = list(big.mark = ","), linesep = "") %>% 
  column_spec(column = 2:8, width = "0.9in") %>% 
    kable_styling(font_size = 8.5, latex_options = "scale_down")
```
\elandscape



\blandscape
### Conviction History

Of those persons who had more than 4 violent arrests in Manhattan while at liberty on a DANY case, below chart provides the average number of their convictions.

```{r, echo = F}
avgCvt %>% 
  kable(format = 'latex', digits = 0, longtable = T,
      booktabs = T, align = c(rep('l', 2), rep('c', 2)),
      format.args = list(big.mark = ","), linesep = "") %>% 
  column_spec(column = 2:8, width = "0.9in") %>% 
    kable_styling(font_size = 9, latex_options = "scale_down")
```



### Gun Cases

Of those persons who had more than 4 violent arrests in Manhattan while at liberty on a DANY case, below chart provides distribution of number of gun cases since 2018.

```{r, echo = F}
gunCases %>% 
  kable(format = 'latex', digits = 0, longtable = T,
      booktabs = T, align = c(rep('l', 2), rep('c', 2)),
      format.args = list(big.mark = ","), linesep = "") %>% 
  column_spec(column = 2:8, width = "0.9in") %>% 
    kable_styling(font_size = 9, latex_options = "scale_down")
```


\elandscape




